\ttitle

### 2015-08-16 Chapters/Introduction.tex
Introduction - Minor changes
- Minor changes to introduction

\subsection{Patient selection}
\subsubsection{Resectability criteria}
Resectable pancreatic cancer is defined as a tumour that does not involve the coeliac axis or the superior mesenteric artery and is not associated with distant metastatic disease. Tumours involving the portal vein or superior mesenteric vein are considered borderline resectable and can still be resected completely (R0) with en-bloc venous resection in selected cases.\parencite{fuhrman_rationale_1996} The role of neoadjuvant therapy such as chemotherapy with or without local radiotherapy and newer treatment modalities such as electroporation\parencite{bower_irreversible_2011}  and intra-peritoneal chemotherapy is the subject of ongoing clinical research in patients with locally advanced and/or metastatic disease. Early reports suggest that some of these patients can undergo potentially curative surgery after neo-adjuvant treatment.

%10/08/2015 - Review completete to here
\subsubsection{Patient factors}

@@

Postoperative pancreatic fistula is one of the most dreaded complications after a pancreaticoduodenectomy and can be associated with significant short-term morbidity as well as long-term disability. The reported incidence of postoperative pancreatic fistula varies from 2\% to 30\% after pancreaticoduodenectomy. \parencite{yeo_six_1997, deoliveira_assessment_2006, bassi_postoperative_2005, winter_biochemical_2007, pratt_risk_2008} The variation in reported incidence has been largely due to lack of clear definition of what constituted a postoperative pancreatic fistula. It can be the result of breakdown or poor healing at the pancreatico-jejunostomy/ pancreatico-gastrostomy or may be the result of direct parenchymal leak unrelated to the anastomosis. It is now generally accepted that 1 in 4 patients will develop a pancreatic fistula as defined by the International Study Group for Pancreatic Fistula (ISGPF) which has published a consensus statement on the definition and grading of postoperative pancreatic fistula.\parencite{bassi_postoperative_2005} A postoperative pancreatic fistula is defined as drain output of any measurable quantity after the third postoperative day with amylase content greater than three times the upper limit of the normal serum amylase value at the laboratory used for testing. Three grades of postoperative pancreatic fistula have been defined based on clinical severity as described in Table \ref{table:isgps_popf} on p\pageref{table:isgps_popf}. Grade B and C fistulae are considered to be clinically significant as they alter patient management and are often associated with other secondary complications such as intra-abdominal sepsis, post-pancreatectomy haemorrhage or delayed gastric emptying. They may require further intervention (either radiological or operative) or prolonged critical care support.

@@

\section{Comorbidity and Risk Stratification}
\label{sec:comorbidity_risk_stratification}
\todo{Comorbidity and Risk Stratification - needs major tidying up - will take several days}
\subsection{Comorbidity}
Comorbidity is defined as the presence of or the effect of other diseases that a patient has in addition to the primary disease of interest. Major pancreatic surgery requires the patient to have adequate reserve to cope not only with the increased physiological demand during and immediately after surgery but also to cope with any adverse events that may occur.

The presence of comorbid conditions is associated with adverse outcomes in patients undergoing treatment for pancreatic cancer\parencite{mann_review_2010} and often limits therapeutic options available due to the associated complications or side effects of surgery or adjuvant treatment. \parencite{sandroussi_sociodemographics_2010}

Patients with multiple comorbidities are more likely to have higher readmission rates, morbidity and mortality following discharge after pancreaticoduodenectomy.\parencite{schneider_patient_2012} DeOliveira and co-workers reported that cardiovascular disease was a risk factor not only for overall morbidity but also complication severity after pancreaticoduodenectomy.\parencite{deoliveira_assessment_2006} Cancer cachexia is associated with increased incidence of complications and mortality after pancreaticoduodenectomy\parencite{pausch_cachexia_2012} while obesity is known to be associated with greater incidence and severity of postoperative complications.\parencite{benns_impact_2009} However, existing methods of measuring the impact of comorbidity on physiological fitness are limited and do not adequately predict outcomes after major pancreatic surgery. \parencite{shah_limitations_2012}

@@

Objective measurement of oxygen delivery at the tissue level at times of physiological stress allows for identification of patients who may struggle during the perioperative phase.

Several tests have been used for preoperative assessment of cardiac function. These include
- electrocardiography
- echocardiography
- exercise tolerance testing
- myocardial perfusion scans

Tests of respiratory function that are commonly performed in selected patients undergoing major surgery include
- pulmonary function tests including forced expiratory volume and forced vital capacity
- spirometry

However, neither of the above cardiac or respiratory function tests adequately measure the ability of the cardiopulmonary and circulatory systems to deliver oxygen to the tissues at times of increased demand.

\section{Cardiopulmonary Exercise Testing}
\todo{A brief history of CPET in general}
\todo{The physiological basis of CPET}

\subsection{Origins and physiological basis of cardiopulmonary exercise testing}
The measurement of aerobic fitness has been an important aspect of improving athletic performance in competitive sports. Cardiopulmonary exercise testing has been used for a long time for detailed measurement of aerobic fitness in athletes as well as offer customised training regimens aimed at improving performance at activities that place a significant stress on the cardiovascular and respiratory systems.

>>>>

\subsection{Resectability criteria}
\label{sec:resectability_criteria}
Resectable pancreatic cancer is defined as a tumour that does not involve the coeliac axis or the superior mesenteric artery and is not associated with distant metastatic disease. Tumours involving the portal vein or superior mesenteric vein are considered borderline resectable and can still be resected completely (R0) with en-bloc venous resection in selected cases.\parencite{fuhrman_rationale_1996} The role of neoadjuvant therapy such as chemotherapy with or without local radiotherapy and newer treatment modalities such as electroporation\parencite{bower_irreversible_2011}  and intra-peritoneal chemotherapy is the subject of ongoing clinical research in patients with locally advanced and/or metastatic disease. Early reports suggest that some of these patients can undergo potentially curative surgery after neo-adjuvant treatment.

%10/08/2015 - Review completete to here

@@

Postoperative pancreatic fistula is one of the most dreaded complications after a pancreaticoduodenectomy and can be associated with significant short-term morbidity as well as long-term disability. The reported incidence of postoperative pancreatic fistula varies from 2\% to 30\% after pancreaticoduodenectomy. \parencite{yeo_six_1997, deoliveira_assessment_2006, bassi_postoperative_2005, winter_biochemical_2007, pratt_risk_2008} The variation in reported incidence has been largely due to lack of clear definition of what constituted a postoperative pancreatic fistula. It can be the result of breakdown or poor healing at the pancreatico-jejunostomy/ pancreatico-gastrostomy or may be the result of direct parenchymal leak unrelated to the anastomosis.

It is now generally accepted that 1 in 4 patients will develop a pancreatic fistula as defined by the International Study Group for Pancreatic Fistula (ISGPF) which has published a consensus statement on the definition and grading of postoperative pancreatic fistula.\parencite{bassi_postoperative_2005} A postoperative pancreatic fistula is defined as drain output of any measurable quantity after the third postoperative day with amylase content greater than three times the upper limit of the normal serum amylase value at the laboratory used for testing.

Three grades of postoperative pancreatic fistula have been defined based on clinical severity as described in Table \ref{table:isgps_popf} on p\pageref{table:isgps_popf}. Grade B and C fistulae are considered to be clinically significant as they alter patient management and are often associated with other secondary complications such as intra-abdominal sepsis, post-pancreatectomy haemorrhage or delayed gastric emptying. They may require further intervention (either radiological or operative) or prolonged critical care support.

@@

\section{Comorbidity and Risk Stratification}
\label{sec:comorbidity_risk_stratification}
\todo{Comorbidity and Risk Stratification - needs major tidying up - will take several days}
Selection of patients suitable for potentially curative surgery not only depends on the resectability criteria that were discussed in section \ref{sec:resectability_criteria} but also on patient fitness. Major pancreatic surgery is associated with significant morbidity. Subjecting patients with poor physiological reserve to a pancreaticoduodenectomy is fraught with risk. It can result in a worse outcome for the patient due to either post-operative mortality or prolonged hospitalisation due to complications resulting in poor quality of life with little difference in longevity as compared to palliation. Patient factors that affect postoperative recovery and long-term survival are therefore important in determining suitability for surgery.

\subsection{Comorbidity}
Comorbidity is defined as the presence of or the effect of other diseases that a patient has in addition to the primary disease of interest. Major pancreatic surgery requires the patient to have adequate reserve to cope not only with the increased physiological demand during and immediately after surgery but also to cope with any adverse events that may occur.

The presence of comorbid conditions is associated with increased incidence of adverse outcomes in patients undergoing treatment for pancreatic cancer\parencite{mann_review_2010}. It can also limit the therapeutic options available due to the associated complications or side effects of surgery or adjuvant treatment. \parencite{sandroussi_sociodemographics_2010} Patients with multiple comorbidities are more likely to have higher readmission rates, morbidity and mortality following discharge after pancreaticoduodenectomy. \parencite{schneider_patient_2012}

DeOliveira and co-workers reported that cardiovascular disease was a risk factor not only for overall morbidity but also complication severity after pancreaticoduodenectomy.\parencite{deoliveira_assessment_2006} Cancer cachexia is associated with increased incidence of complications and mortality after pancreaticoduodenectomy \parencite{pausch_cachexia_2012} while obesity is known to be associated with greater incidence and severity of postoperative complications.\parencite{benns_impact_2009} However, existing methods of measuring the impact of comorbidity on physiological fitness are limited and do not adequately predict outcomes after major pancreatic surgery. \parencite{shah_limitations_2012}

@@

Conventional tests of cardiac and respiratory function assess either performance at rest or only one component of the cardiorespiratory system. These tests include electrocardiography, echocardiography, exercise tolerance testing, myocardial perfusion scan, pulmonary function tests including spirometry. However, none of these tests adequately measure the ability of the cardiopulmonary and circulatory systems to deliver oxygen to the tissues at times of increased demand. Patients with subclinical cardiopulmonary dysfunction or limitation cannot be identified on these tests alone and it is this group of patients who are most likely to be at greater risk of complications and who will most benefit from targeted prehabilitation. Cardiopulmonary exercise testing overcomes the limitations of the conventional static tests by providing a global assessment of the patient's oxygen delivery mechanisms at times of increased physiological stress and demand.

One of the earliest attempts to develop a dynamic test in surgical patients was reported by Starr and co-workers in 1945 who tried to devise a test that could assess respiratory function and oxygen utilisation in surgical patients before, immediately after and a week after surgery. They reported on the difficulties in obtaining reproducible results but suggested that a slow return to baseline of heart rate, respiratory rate and oxygen utilisation after surgery may indicate patients who are slow to recover. \parencite{starr_convalescence_1945}

\section{Cardiopulmonary Exercise Testing}
%Not sure I need these two sections - delete once done vacillating
\todo{A brief history of CPET in general}
\todo{The physiological basis of CPET}

\subsection{Origins and physiological basis of cardiopulmonary exercise testing}
The measurement of aerobic fitness has been an important aspect of improving athletic performance in competitive sports. Cardiopulmonary exercise testing has been used for a long time for detailed measurement of aerobic fitness in athletes. It has also been used to design customised training regimens aimed at improving performance at activities that place a significant stress on the cardiovascular and respiratory systems by allowing the optimisation of the individual components of the oxygen delivery system.
